Traffic Collisions in Toronto

### 2024-09-27 paper/paper.qmd
polish paper

---
title: "Traffic Collisions in Toronto: Analysis of Key Variables"
subtitle: "An Analysis of Traffic Collision Data from 2014-2024"
author: 
  - Andy Jiang
thanks: "Code and data are available at: https://github.com/AndyYanxunJiang/toronto-traffic-collisions"
date: today
date-format: long
abstract: "This paper analyzes traffic collisions in Toronto from 2014 to 2024, focusing on patterns across months, days of the week, times of day, and types of collisions, including injuries, property damage, and failed to remain incidents. Fatalities are analyzed separately as a distinct trend. The data, sourced from Open Data Toronto, reveals a light W-shaped trend in monthly collisions, with peaks in January and June. Collisions are most frequent between 3 PM and 5 PM, with Friday being the highest day for collisions and Sunday the lowest. Property damage accounts for about 71% of all incidents. The number of fatalities has generally declined over the years, though trends fluctuate annually. These findings highlight the need for ongoing road safety initiatives and targeted interventions during high-risk periods to improve long-term traffic safety."
format: pdf
bibliography: references.bib
---

@@

@tbl-1 provides a random sample of the cleaned dataset, displaying all relevant variables used for analysis:

```{r}
#| label: tbl-1
#| tbl-cap: Random sample of the cleaned collision data
#| echo: false

@@

```{r}
#| label: fig-collisions-day
#| fig-cap: Total traffic collisions by day of the week from 2014 to 2024
#| echo: false

@@

```{r}
#| label: fig-collisions-month
#| fig-cap: Monthly traffic collision counts in Toronto from 2014 to 2024.
#| echo: false

@@

```{r}
#| label: fig-collisions-hour
#| fig-cap: Total collisions by hour of the day from 2014 to 2024
#| echo: false

@@

```{r}
#| label: fig-collisions-type
#| fig-cap: Proportion of each type of traffic collisions (2014–2024).
#| echo: false

@@

```{r}
#| label: fig-fatalities
#| fig-cap: Yearly trend in fatalities from collisions (2014-2024)
#| echo: false

@@

Sunday, with the lowest number of collisions, likely benefits from reduced road congestion, as many people tend to stay home or partake in local activities that don’t require long-distance travel. This also suggests that weekends, despite the occasional recreational travel, generally have safer road conditions. It is worth considering whether policies or public safety campaigns targeted at reducing Friday incidents could be beneficial, perhaps through encouraging more cautious driving during this high-risk period.

@@

## Collision Types Implications
The distribution of collision types adds depth to our understanding of how most accidents unfold. The dominance of property damage collisions—representing approximately 71% of all incidents—suggests that the majority of accidents are minor, involving either physical contact between vehicles or collisions with roadside infrastructure like signs or barriers. These accidents often result in fender benders or minor vehicle damage but don’t necessarily lead to injuries or fatalities.

>>>>

---
title: "Traffic Collisions in Toronto"
subtitle: "Analysis of Trends and Safety Insights"
author: 
  - Andy Jiang
thanks: "Code and data are available at: https://github.com/AndyYanxunJiang/toronto-traffic-collisions."
date: today
date-format: long
abstract: "This paper analyzes traffic collisions in Toronto from 2014 to 2024, focusing on patterns across months, days of the week, times of day, and types of collisions, including injuries, property damage, and failed to remain incidents. Fatalities are analyzed separately as a distinct trend. The data, sourced from Open Data Toronto, reveals a light W-shaped trend in monthly collisions, with peaks in January and June. Collisions are most frequent between 3 PM and 5 PM, with Friday being the highest day for collisions and Sunday the lowest. Property damage accounts for about 71% of all incidents. The number of fatalities has generally declined over the years, though trends fluctuate annually. These findings highlight the need for ongoing road safety initiatives and targeted interventions during high-risk periods to improve long-term traffic safety."
format: pdf
bibliography: references.bib
---

@@

@tbl-sample provides a random sample of the cleaned dataset, displaying all relevant variables used for analysis:

```{r}
#| label: tbl-sample
#| tbl-cap: Random sample of 10 rows from the cleaned collision data
#| echo: false

@@

```{r}
#| label: fig-collisions-day
#| fig-cap: Total traffic collisions by day of the week in Toronto from 2014 to 2024.
#| echo: false

@@

```{r}
#| label: fig-collisions-month
#| fig-cap: Total traffic collisions by month in Toronto from 2014 to 2024.
#| echo: false

@@

```{r}
#| label: fig-collisions-hour
#| fig-cap: Total traffic collisions by hour of the day in Toronto from 2014 to 2024.
#| echo: false

@@

```{r}
#| label: fig-collisions-type
#| fig-cap: Proportion of each type of traffic collisions in Toronto from 2014 to 2024.
#| echo: false

@@

```{r}
#| label: fig-fatalities
#| fig-cap: Yearly trend in fatalities from traffic collisions in Toronto from 2014 to 2024.
#| echo: false

@@

Sunday, with the lowest number of collisions, likely benefits from reduced road congestion, as many people tend to stay home or partake in local activities that don't require long-distance travel. This also suggests that weekends, despite the occasional recreational travel, generally have safer road conditions. It is worth considering whether policies or public safety campaigns targeted at reducing Friday incidents could be beneficial, perhaps through encouraging more cautious driving during this high-risk period.

@@

## Collision Types Implications
The distribution of collision types adds depth to our understanding of how most accidents unfold. The dominance of property damage collisions—representing approximately 71% of all incidents—suggests that the majority of accidents are minor, involving either physical contact between vehicles or collisions with roadside infrastructure like signs or barriers. These accidents often result in fender benders or minor vehicle damage but don't necessarily lead to injuries or fatalities.
